A Large-Scale Study of the Use of Eval in R

### 2020-11-05 inst/paper/main.tex
Reorganize sections

Traditional program analysis techniques are based on sound reasoning about
program behavior, a program analyzer computes an over-approximation of the
set of possible operations performed by the system being
analyzed~\cite{cc77}.  This is predicated on knowing the program's code, and
computing that code's behavior for all possible inputs. The presence of
\eval entails that the program's input, any string variable, may be
reflected into its code giving rise to \emph{any} behavior allowed by the
language semantics. In many dynamic languages, \eval can redefine most
user-defined and built-in functions resulting in a complete loss of
precision for any operation following \eval. A group of influential
resaerchers argued to give up on soundness and, instead, to
under-approximate dynamic features~\cite{soundy} In their words ``a
practical analysis, therefore, may pretend that \eval does nothing, unless
it can precisely resolve its string argument at compile time.''  Assuming
that \eval does not have side-effects may be too optimistic. We rather
propose to study what \eval does in practice, in real-world code, so as to
give researchers data about how to better under-approximate uses of that
feature.

@@

\section{Infrastructure}

@@

\section{Corpus}

@@

\subsection{CRAN Packages}

@@

\subsection{Kaggle Scripts}

@@

\section{Usage Metrics}

@@

\section{A taxonomy of eval}

@@

\subsection{Operation Mix}

@@

\subsection{Provenance}

\subsection{Side-Effects}

\subsection{Scope}

\subsection{Patterns}

@@

\subsection{Consistency}

We look at how \emph{consistent} the \c{expr} argument of \eval can be, \ie
how many different types of the resolved \c{expr} there are per call
sites. Most of the call sites, \ie \PercentMonomorphic, are
\emph{consistent}, and this is similar to javascript. However, a few ones
are highly \emph{polymorphic} (10 different types). They are the pipe
operators \c{\%>\%}, \c{\%<>\%} and \c{\%$\%} in package \emph{magrittr}. It
is effectively used to compose functions on their first argument, which can
be of any type.

\section{Threats to Validity} The primary threat to validity is the issue of
code coverage, a cause of concern for any dynamic approach. We mitigate this
by including focusing on R packages that have high code coverage. We turn
off the bytecode compiler for this study. The bytecode compiler can also
call \eval. We do not get source locations for \UndefinedEvalsRnd eval
calls. In these cases \eval is either passed as an argument to a
higher-order functions or is defined in a function returned by a
higher-order function and the R parser does not retain location information
for \eval. However, this is a meager \PercentUndefinedEval of all eval calls
and is unlikely to affect our analysis. We ignore calls to the native \eval
function exposed by R. We also ignore the \c{rlang::tidy_eval} function
which uses native \eval internally because \c{rlang} is used to implement a
DSL for data analysis in R. It introduces a new first-class promise object
called \c{quosure} for which it implements special evaluation support in
\c{tidy_eval}.

>>>>

Traditional program analysis techniques are based on sound reasoning about
program behavior, a program analyzer computes an over-approximation of the
set of possible operations performed by the system being
analyzed~\cite{cc77}.  This is predicated on knowing the program's code, and
computing that code's behavior for all possible inputs. The presence of
\eval entails that the program's input, any string variable, may be
reflected into its code giving rise to \emph{any} behavior allowed by the
language semantics. In many dynamic languages, \eval can redefine most
user-defined and built-in functions resulting in a complete loss of
precision for any operation following \eval. A group of influential
researchers argued to give up on soundness and, instead, to
under-approximate dynamic features~\cite{soundy} In their words ``a
practical analysis, therefore, may pretend that \eval does nothing, unless
it can precisely resolve its string argument at compile time.''  Assuming
that \eval does not have side-effects may be too optimistic. We rather
propose to study what \eval does in practice, in real-world code, so as to
give researchers data about how to better under-approximate uses of that
feature.

@@

\section{Methodology}

\subsection{Infrastructure}

@@

\subsection{Corpus}

@@

\subsubsection{CRAN Packages}

@@

\subsubsection{Kaggle Scripts}

@@

\subsection{Threats to Validity}
The primary threat to validity is the issue of
code coverage, a cause of concern for any dynamic approach. We mitigate this
by including focusing on R packages that have high code coverage. We turn
off the bytecode compiler for this study. The bytecode compiler can also
call \eval. We do not get source locations for \UndefinedEvalsRnd eval
calls. In these cases \eval is either passed as an argument to a
higher-order functions or is defined in a function returned by a
higher-order function and the R parser does not retain location information
for \eval. However, this is a meager \PercentUndefinedEval of all eval calls
and is unlikely to affect our analysis. We ignore calls to the native \eval
function exposed by R. We also ignore the \c{rlang::tidy_eval} function
which uses native \eval internally because \c{rlang} is used to implement a
DSL for data analysis in R. It introduces a new first-class promise object
called \c{quosure} for which it implements special evaluation support in
\c{tidy_eval}.

\section{Analysis of Eval}

This section presents the results of our empirical study of \eval in R as
answers to four research questions.

\subsection{Use of Eval}

The first research question we address is how and why \eval is used in R code.

\begin{center}{\bfseries RQ1: What are the patterns of usage of \eval in the corpus?}\end{center}

@@

\subsubsection{A taxonomy of eval}

@@

\subsubsection{Operation Mix}

@@

\subsubsection{Usage Patterns}

@@

\subsection{Comparison with eval in Javascript}

\subsubsection{Consistency}

We look at how \emph{consistent} the \c{expr} argument of \eval can be, \ie
how many different types of the resolved \c{expr} there are per call
sites. Most of the call sites, \ie \PercentMonomorphic, are
\emph{consistent}, and this is similar to javascript. However, a few ones
are highly \emph{polymorphic} (10 different types). They are the pipe
operators \c{\%>\%}, \c{\%<>\%} and \c{\%\$\%} in package \emph{magrittr}. It
is effectively used to compose functions on their first argument, which can
be of any type.

\subsection{How Dangerous is eval}

\subsubsection{Provenance}

\subsubsection{Side-Effects}

\subsubsection{Scope}

\subsection{Removing Eval}
